---
title: "CEMA INTERNSHIP"
author: "Bonsoul"
date: "2025-04-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,libraries}



```

```{r,fh}
library(tidyverse)
library(glmmTMB)
library(ggplot2)
library(readr)
library(stringr)
library(Amelia)
library(lme4)
library(Matrix)
library(knitr)
library(scales)
library(sf)

```

```{r,load the dataset}
df <- read.csv("Data/HIV data 2000-2023.csv")

```

```{r, EDA}
str(df)

dim(df)

summary(df)

colSums(is.na(df))
```

# the dataset is clean. Will rename two columns Period and Location

```{r, rename columns}
df1 <- df %>%
  rename(
    Country = Location,
    Year = Period
  )
```

```{r, extract the first number and remove spaces}
df1$Value_clean <- str_extract(df1$Value,"^[0-9 ]+") %>%
  str_replace_all(" ", "") %>%
  as.numeric()

View(df1)
```

```{r}
# Filter for Kenya and a specific indicator
df_kenya <- df1 %>%
  filter(Country == "Kenya", 
         Indicator == "Estimated number of people (all ages) living with HIV")

# Plot
ggplot(df_kenya, aes(x = Year, y = Value_clean)) +
  geom_line(color = "red", size = 1.2) +
  geom_point() +
  labs(title = "HIV Estimates in Kenya (2000–2023)",
       y = "Estimated People Living with HIV",
       x = "Year") +
  theme_minimal()
```

```{r,global top 75% countries}
df_top75_global <- df1 %>%
  group_by(Country) %>%
  mutate(total = sum(Value_clean, na.rm = TRUE)) %>%
  ungroup() %>%
  arrange(desc(total)) %>%
  mutate(cum_perc = cumsum(total) / sum(total)) %>%
  filter(cum_perc <= 0.75)

ggplot(df_top75_global, aes(x = Year, y = Value_clean, color = Country)) +
  geom_line(size = 1) +
  labs(title = "HIV Trends in Countries Contributing to 75% of Global Burden",
       y = "Estimated Cases", x = "Year") +
  theme_minimal()
```


```{r,WHO 75% region}
# First, group and filter within each WHO region
df_top75_by_region <- df1 %>%
  group_by(ParentLocationCode, Country) %>%
  mutate(total = sum(Value_clean, na.rm = TRUE)) %>%
  arrange(ParentLocationCode, desc(total)) %>%
  group_by(ParentLocationCode) %>%
  mutate(cum_perc = cumsum(total) / sum(total)) %>%
  filter(cum_perc <= 0.75)

ggplot(df_top75_by_region, aes(x = Year, y = Value_clean, color = Country)) +
  geom_line(size = 1.1) +
  facet_wrap(~ ParentLocationCode, scales = "free_y") +
  labs(
    title = "HIV Trends in Countries Contributing to 75% of the Burden Within Each WHO Region",
    subtitle = "Different scales per region highlight within-region trends",
    y = "Estimated HIV Cases",
    x = "Year",
    caption = "Source: WHO HIV Data"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 20, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 14, color = "gray40"),
    plot.caption = element_text(size = 10, hjust = 1, color = "gray50"),
    strip.text = element_text(size = 14, face = "bold"),  # Bigger facet labels
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 12),
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = 10),
    legend.key.width = unit(1.5, "cm"),
    legend.key.height = unit(0.5, "cm"),
    legend.box = "vertical"
  ) +
  scale_y_continuous(labels = comma) +
  guides(color = guide_legend(ncol = 3))

```

```{r, multidimensional poverty headcount data}

data <- read.csv("~/R Studio/R_Statistical-Language/Data/multidimensional_poverty.csv")
```

```{r,data cleaning}
# Check for NA values and sum per column
null_counts <- colSums(is.na(data))

# View the result# Convert to a data frame for nicer display
null_counts_df <- data.frame(Column = names(null_counts), Null_Count = null_counts)

# View the result in a kable
kable(null_counts_df, caption = "Null Counts per Column")

```


# Most columns, including Region, Country.Code, Economy, Reporting.Year, Survey.Name, Survey.Year, Survey.Coverage, Welfare.Type, Survey.Comparability, Monetary, and Poverty.Headcount.Ratio, have no missing values. However, some key indicators show gaps: Educational Enrollment has the highest number of missing values (36), followed by Sanitation (25), Drinking Water (9), and minor gaps in Educational Attainment and Electricity (1 each). These missing values, especially in education and infrastructure, may impact the completeness of subsequent analyses.


```{r, fill the nulls}
data1 <- data %>%
  mutate_all(~ ifelse(is.na(.), 0, .))
```

```{r,colnames}

colnames(df1)

colnames(data1)

data2 <- data1 %>%
  rename(
    Country = Economy
  )

colnames(data2)
```

```{r, join}
merged_df <- data2 %>%
  left_join(df1, by = c("Country" = "Country"))
```


```{r, cleaning the data}
missmap(merged_df, main = "Missing values heatmap", col = c("red", "grey"))


merged_df_clean <- merged_df %>%
  mutate(
    Value_clean = ifelse(is.na(Value_clean), 0, Value_clean)
  )

```

```{r,distribution}


# Poverty Headcount Ratio
ggplot(merged_df_clean, aes(x = Poverty.Headcount.Ratio)) +
  geom_histogram(bins = 30, fill = "darkorange") +
  theme_minimal()

# 2. Poverty Headcount Ratio (Income)
ggplot(merged_df_clean, aes(x = Poverty.Headcount.Ratio)) +
  geom_histogram(bins = 30, fill = "darkorange") +
  theme_minimal() +
  labs(title = "Histogram of Poverty Headcount Ratio", x = "Poverty Headcount Ratio", y = "Count")

# 3. Educational Attainment
ggplot(merged_df_clean, aes(x = Educational.Attainment)) +
  geom_histogram(bins = 30, fill = "forestgreen") +
  theme_minimal() +
  labs(title = "Histogram of Educational Attainment", x = "Educational Attainment", y = "Count")

# 4. Educational Enrollment (School Enrollment)
ggplot(merged_df_clean, aes(x = Educational.Enrollment)) +
  geom_histogram(bins = 30, fill = "purple") +
  theme_minimal() +
  labs(title = "Histogram of Educational Enrollment", x = "Educational Enrollment", y = "Count")

# 5. Electricity Access
ggplot(merged_df_clean, aes(x = Electricity)) +
  geom_histogram(bins = 30, fill = "dodgerblue") +
  theme_minimal() +
  labs(title = "Histogram of Electricity Access", x = "Electricity Access", y = "Count")

# 6. Sanitation
ggplot(merged_df_clean, aes(x = Sanitation)) +
  geom_histogram(bins = 30, fill = "tomato") +
  theme_minimal() +
  labs(title = "Histogram of Sanitation", x = "Sanitation", y = "Count")

# 7. Drinking Water
ggplot(merged_df_clean, aes(x = Drinking.Water)) +
  geom_histogram(bins = 30, fill = "goldenrod") +
  theme_minimal() +
  labs(title = "Histogram of Drinking Water Access", x = "Drinking Water", y = "Count")
```

```{r,}
# Select numeric columns only
num_data <- merged_df_clean %>%
  select_if(is.numeric)

# Correlation matrix
cor_matrix <- cor(num_data, use = "complete.obs")

# Visualize
library(corrplot)
corrplot(cor_matrix, method = "color", tl.cex = 0.8)

```


```{r}
# HIV vs. Poverty Headcount
ggplot(merged_df_clean, aes(x = Poverty.Headcount.Ratio, y = Value_clean)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", color = "red") +
  theme_minimal()

# HIV vs. Educational Enrollment
ggplot(merged_df, aes(x = Educational.Enrollment, y = Value_clean)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", color = "green") +
  theme_minimal()

```


```{r}
# Average HIV Prevalence by Region (if available)
merged_df_clean %>%
  group_by(Region) %>%
  summarise(Avg_HIV = mean(Value_clean, na.rm = TRUE),
            Avg_Poverty = mean(Poverty.Headcount.Ratio, na.rm = TRUE)) %>%
  arrange(desc(Avg_HIV))

```


```{r,summary}
summary(merged_df_clean[, c("Value_clean", 
                             "Poverty.Headcount.Ratio", 
                             "Educational.Attainment", 
                             "Educational.Enrollment", 
                             "Electricity", 
                             "Sanitation", 
                             "Drinking.Water")])

```

```{r}
merged_df_clean2 <- merged_df_clean %>%
  mutate(
    Electricity_binary = ifelse(Electricity > 0, 1, 0),
    Sanitation_binary = ifelse(Sanitation > 0, 1, 0),
    Drinking_Water_binary = ifelse(Drinking.Water > 0, 1, 0)
  )



```

```{r,model}
merged_df_clean3 <- merged_df_clean2 %>%
  mutate(across(c(Poverty.Headcount.Ratio, Educational.Attainment, Educational.Enrollment), scale))

merged_df_clean3$Value_clean_log <- log1p(merged_df_clean3$Value_clean)

model2 <- glmmTMB(Value_clean_log ~ Poverty.Headcount.Ratio + Educational.Attainment + Educational.Enrollment +
                    (1 | Country.Code) + (1 | Reporting.Year),
                  data = merged_df_clean,
                  control = glmmTMBControl(optCtrl = list(iter.max = 10000, eval.max = 10000)))

```


```{r, Mortality dataset}
mortality_df <- read.csv("Data/dataset_datascience.csv")

shapefile <- st_read("EA Shape Files/ESA_admin1_region.shp")
```

```{r,eac countries}
eac_countries <- c("Burundi", "Kenya", "Rwanda", "South Sudan", 
                   "Tanzania", "Uganda", "Democratic Republic of the Congo", "Somalia")
```

```{r,# 🧹 Filter data for EAC countries only}

mortalityeac_df <- mortality_df %>% filter(Geographic.area %in% eac_countries)
```

```{r,check}
unique(shapefile$COUNTRY)

```

```{r}
# List of East African Community countries
eac_countries <- c("Burundi", "Kenya", "Rwanda", "South Sudan", 
                   "Tanzania", "Uganda","Somalia","Congo DRC")


# Filter shapefile for only EAC countries
eac_shapefile <- shapefile %>%
  filter(COUNTRY %in% eac_countries)

# Check the result
plot(st_geometry(eac_shapefile))
```


```{r, change Congo DR to Democratic Republic of the Congo }
eac_shapefile <- eac_shapefile %>%
  mutate(COUNTRY = ifelse(COUNTRY == "Congo DRC", "Democratic Republic of the Congo", COUNTRY))

```

```{r, merge}
colnames(mortality_df)

latest_year <- max(mortality_df$Series.Year, na.rm = TRUE)

```

```{r, summarise}
# Step 1: Fix shapefile
eac_shapefile <- st_make_valid(eac_shapefile)

shapefile_country <- eac_shapefile %>%
  group_by(COUNTRY) %>%
  summarize(geometry = st_union(geometry)) %>%
  ungroup()
```

```{r, latest mortality rate}
# Prepare latest under-five mortality data
under5_data <- mortality_df %>%
  filter(Indicator == "Under-five mortality rate", Geographic.area %in% eac_countries) %>%
  select(Country = Geographic.area, Series.Year)
```

```{r, latest mortality}
# Prepare latest neonatal mortality data
neonatal_data <- mortality_df %>%
  filter(Indicator == "Neonatal mortality rate", Geographic.area %in% eac_countries) %>%
  select(Country = Geographic.area, Series.Year)

```

```{r,latest}

```
```{r,}
View(neonatal_data)
```

```{r,merge}
shape_merged <- shapefile_country %>%
  filter(COUNTRY %in% eac_countries) %>%
  left_join(under5_data, by = c("COUNTRY" = "Country")) %>%
  left_join(neonatal_data, by = c("COUNTRY" = "Country"))
```

```{r, view}
View(shape_merged)
```

```{r,plot}
ggplot(shape_merged) +
  geom_sf(aes(fill = Neonatal_Mortality)) +
  scale_fill_gradient(name = "Neonatal Mortality\n(per 1,000)", low = "lightblue", high = "darkblue") +
  theme_minimal() +
  ggtitle("Latest Neonatal Mortality Rate in East African Community Countries")
```