---
title: "HIV/AIDS Modelling"
author: "Bonsoul"
date: "2025-02-26"
output: html_document
---


```{r setup, include=TRUE , warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)  

 

library(rnaturalearth)
library(rnaturalearthdata)
library(CARBayes)
library(gridExtra)             
library(grid) 
library(survival)
library(tidyverse) 
library(caret) 
library(randomForest) 
library(ggplot2) 
library(dplyr) 
library(corrplot) 
library(readxl) 
library(tidyr) 
library(survminer) 
library(sf) 
library(tmap)
library(RColorBrewer)
library(knitr)
library(mgcv)
library(knitr)
library(kableExtra)
library(broom)
library(spdep)
library(brglm2)
library(caret)
```

```{r, Readind the data , warning=FALSE}

df <- read_excel("D:/Downloads/HIV - VARIABLES.xlsx")

str(df)
```

```{r, Preprocessing}
#Rename coordinate columns for clarity
#df <- df %>% rename(latitude = var52, longitude = var51)

duplicates <- df[duplicated(df),] #checking duplicates

print(duplicates)


missing_per_column <- colSums(is.na(df)) #columns with null values
total_missing <- sum(missing_per_column) 
print(total_missing) 
print(missing_per_column)

columns_with_missing <- names(missing_per_column[missing_per_column > 0]) 
print(columns_with_missing)

```

```{r, descriptive analysis}
#Compute summary statistics for numeric columns
summary_stats <- df %>% 
  select(where(is.numeric)) %>% 
  summarise(across(
    everything(), 
    list(
      min = ~min(., na.rm = TRUE),
      Q1 = ~quantile(., 0.25, na.rm = TRUE),
      median = ~median(., na.rm = TRUE),
      mean = ~mean(., na.rm = TRUE),
      Q3 = ~quantile(., 0.75, na.rm = TRUE),
      max = ~max(., na.rm = TRUE),
      sd = ~sd(., na.rm = TRUE)
    )
  ))
summary(summary_stats)

```

```{r,hitograms}
#Visualize distributions with histograms

ggplot(df, aes(x = age)) + 
  geom_histogram(binwidth = 5, fill = "blue", alpha = 0.7) + 
  ggtitle("Histogram of Age") + 
  theme_minimal()
  
  
ggplot(df, aes(x = weightkg)) + 
  geom_histogram(binwidth = 5, fill = "blue", alpha = 0.7) + 
  ggtitle("Histogram of weightkg") + 
  theme_minimal()
  

ggplot(df, aes(x = durationofbfmonths)) + 
  geom_histogram(binwidth = 5, fill = "blue", alpha = 0.7) + 
  ggtitle("Histogram of durationofbfmonths") + 
  theme_minimal()
```

```{r,Boxplots}
for (col in numeric_cols) { 
p <- ggplot(df, aes(y = .data[[col]])) +  
    geom_boxplot(fill = "lightblue", na.rm = TRUE) + 
    ggtitle(paste("Boxplot of", col)) + 
    theme_minimal() 
    print(p)  
}
```

```{r,Mutating +ve and -ne to be numerical}
df1 <- df %>% mutate(monthsantibodytest = ifelse(tolower(monthsantibodytest) == "positive", 1, 0))

table(df1$monthsantibodytest)
```


```{r,survival analysis}
#Create survival object
surv_obj <- Surv(time = df1$time, event = df1$monthsantibodytest)

#Fit Kaplan-Meier survival model
km_fit <- survfit(surv_obj ~ 1, data = df1)

#Plot Kaplan-Meier curve
ggsurvplot(km_fit, data = df1, conf.int = TRUE, ggtheme = theme_minimal(), title = "Kaplan-Meier Survival Curve", xlab = "Time (Months)", ylab = "HIV-Free Survival Probability")

```


```{=html}
<!--Chi-square test used for comparisons
-->
```

colnames(df1)

dependent_var <- "monthsantibodytest"

categorical_vars <- c(
  "subcounty", "maritalstatus", "historyofstiduringpregnancy", "hypertension", 
  "partnershivstatus", "whohivdiseasestage", "t_membrane_r", "sexofthebaby", 
  "residence", "syphillis", "cd4cellcountcellsmm3", "employmentstatus", 
  "ancattendance", "facilitylevelanc", "hivstatusbeforepregnancy", "syphilis", 
  "uti", "miscarriage", "abortion", "stillbirth", "premature", 
  "haartduringpregnancy", "cd4cellcount", "historysti_during_pregnancy", "distancetofacility",
  "genitalwart", "vaginaldischargesyndrome", "genitalherpes", 
  "treatedduringpregnancy", "malaria", "anaemia", "hypertension", 
  "partnerhivstatus", "whohivdiseasestage", "t_membrane_r", "sexbaby", 
  "educationlevel", "adherence","tmembraner", "hypertention","patnershivstatus"
)



all_vars <- categorical_vars[categorical_vars %in% names(df1)]


chi_results <- lapply(all_vars, function(var) {
  table <- table(df[[var]], df[[dependent_var]])
  test <- chisq.test(table)
  data.frame(variable = var, chi2_value = test$statistic, p_value = test$p.value)
})

# Combine the results into a single data frame
chi_result_df <- do.call(rbind, chi_results)

kable(chi_result_df, format = "html", digits = 4, caption = "Table 1: Chi-square Test Results for HIV Antigen Positivity") %>%
  kable_styling(full_width = FALSE, position = "left", font_size = 14) %>%
  column_spec(1, bold = TRUE) %>%   
  column_spec(2, width = "5cm") %>% 
  column_spec(3, width = "5cm") %>% 
  footnote(general = "Source: Computed results based on chi-square test.")





<!--  Factors Associated with HIV Antigen Positivity by Bivariate Logistic Regression-->


dependent_var <- "monthsantibodytest"  

# List of independent variables for bivariate logistic regression
independent_vars <- c(
  "ancattendance","haartduringpregnancy", "whohivdiseasestage","numberofancvisitsmade","lduration","durationofbfmonths","educationlevel","adherence","hivstatusbeforepregnancy","premature") 

# Filter variables that exist in df
available_vars <- independent_vars[independent_vars %in% names(df1)]



# Function to perform bivariate logistic regression
logit_results <- lapply(available_vars, function(var) {
  formula <- as.formula(paste(dependent_var, "~", var))
  model <- glm(formula, data = df1, family = binomial)
  tidy_model <- tidy(model, conf.int = TRUE)  # Get summary with confidence intervals
  
  # Format output
  data.frame(
    Variable = var,
    Odds_Ratio = round(exp(tidy_model$estimate[2]), 4),  # Exponentiate coefficient
    CI = paste0("[", round(exp(tidy_model$conf.low[2]), 3), "-", round(exp(tidy_model$conf.high[2]), 3), "]"),
    P_Value = round(tidy_model$p.value[2], 4)
  )
})

bivar_logit_results_df <- do.call(rbind, logit_results)


kable(bivar_logit_results_df, format = "html", digits = 4, caption = "Table 4.2: Factors Associated with HIV Antigen Positivity by Bivariate Logistic Regression") %>%
  kable_styling(full_width = FALSE, position = "left", font_size = 14) %>%
  column_spec(1, bold = TRUE) %>%   
  column_spec(2, width = "5cm") %>% 
  column_spec(3, width = "5cm") %>% 
  column_spec(4, width = "5cm") %>% 
  footnote(general = "Source: Computed results based on bivariate logistic regression.")



<!--Factors Associated With HIV Antigen Positivity by Multivariate Logistic Regression-->

# Define and run the multivariate logistic regression model
multivar_model <- glm(
  monthsantibodytest ~ lduration + durationofbfmonths + 
    ancattendance + haartduringpregnancy + hivstatusbeforepregnancy + educationlevel + 
    adherence + numberofancvisitsmade + premature,
  data = df1, 
  family = binomial
)

# Extract results
multivar_results <- tidy(multivar_model, conf.int = TRUE) %>%
  filter(term != "(Intercept)") %>%
  mutate(
    Variable = recode(term, 
                      "duration_labor" = "Duration of labor",
                      "duration_breastfeeding" = "Duration of breast feeding in months",
                      "antenatal_clinic_attendance" = "Antenatal clinic attendance",
                      "haart_during_pregnancy" = "Initiation of HAART during pregnancy",
                      "never_attended_school" = "Never attended school",
                      "attended_primary" = "Attended up to Primary school",
                      "attended_secondary" = "Attended up to Secondary school",
                      "adherencepoor" = "Adherence to HAART during ANC (Poor)",
                      "adherencegood" = "Adherence to HAART during ANC (Good)",
                      "no_anc_visits" = "No of ANC visits made during pregnancy",
                      "premature" = "Premature"),
    Odds_Ratio = round(exp(estimate), 4),
    CI = paste0("[", round(exp(conf.low), 3), "-", round(exp(conf.high), 3), "]"),
    P_Value = round(p.value, 4)
  ) %>%
  select(Variable, Odds_Ratio, CI, P_Value)

# Display the results in a table

kable(multivar_results, format = "html", digits = 4, 
      caption = "Table 4.3: Factors Associated With HIV Antigen Positivity by Multivariate Logistic Regression") %>%
  kable_styling(full_width = FALSE, position = "left", font_size = 14) %>%
  column_spec(1, bold = TRUE) %>%  
  column_spec(2:4, width = "5cm") %>%
  footnote(general = "Source: Computed results based on multivariate logistic regression analysis.")



<!--Bivariate Cox PH Model-->

# List of independent variables
independent_vars <- c("employmentstatus", "ancattendance", "facilitylevelanc",
                      "hivstatusbeforepregnancy", "premature","haartduringpregnancy",
                      "historyofstiduringpregnancy", "cd4cellcountcellsmm3", "whohivdiseasestage",
                      "numberofancvisitsmade", "durationofbfmonths")

# Apply Bivariate Cox PH Models for each variable

cox_results <- lapply(independent_vars, function(var) {
  formula <- as.formula(paste("Surv(time, monthsantibodytest) ~", var))
  model <- coxph(formula, data = df1)
  tidy_model <- tidy(model, conf.int = TRUE)  # Extract HR, CI, p-value
  
  # Format the output
  data.frame(
    Variable = var,
    HR = round(exp(tidy_model$estimate), 3),  # Exponentiate to get Hazard Ratio
    CI = paste0("[", round(exp(tidy_model$conf.low), 3), "-", round(exp(tidy_model$conf.high), 3), "]"),
    P_Value = round(tidy_model$p.value, 3)
  )
})

# Combine results into a single table
cox_results_df <- bind_rows(cox_results)


# Generate Harvard-style table with kableExtra
kable(cox_results_df, format = "html", digits = 4, caption = "Table 4.4: Factors Associated With HIV Antigen Positivity by Bivariate Cox PH Model") %>%
  kable_styling(full_width = FALSE, position = "left", font_size = 14) %>%
  column_spec(1, bold = TRUE) %>%   # Make first column (Variable) bold
  column_spec(2, width = "5cm") %>% # Adjust HR column width
  column_spec(3, width = "6cm") %>% # Adjust CI column width
  column_spec(4, width = "4cm") %>% # Adjust P-value column width
  footnote(general = "Source: Computed results based on statistical analysis.")

<!--Multivariate-->

colnames(df1)

# List of independent variables
independent_vars <- c("hivstatusbeforepregnancy","haartduringpregnancy","cd4cellcountcellsmm3", "whohivdiseasestage","adherence")

# Create a formula with all independent variables for the multivariate model
multivariate_formula <- as.formula(paste("Surv(time, monthsantibodytest) ~", paste(independent_vars, collapse = " + ")))

# Apply Multivariate Cox PH Model
multivariate_model <- coxph(multivariate_formula, data = df1)

# Tidy the model to extract HR, CI, and p-values
tidy_multivariate_model <- tidy(multivariate_model, conf.int = TRUE)

# Format the output
multivariate_results_df <- data.frame(
  Variable = tidy_multivariate_model$term,
  HR = round(exp(tidy_multivariate_model$estimate), 3),  # Exponentiate to get Hazard Ratio
  CI = paste0("[", round(exp(tidy_multivariate_model$conf.low), 3), "-", round(exp(tidy_multivariate_model$conf.high), 3), "]"),
  P_Value = round(tidy_multivariate_model$p.value, 3)
)

# Generate Harvard-style table with kableExtra
kable(multivariate_results_df, format = "html", digits = 4, caption = "Table 4.X: Factors Associated With HIV Antigen Positivity by Multivariate Cox PH Model") %>%
  kable_styling(full_width = FALSE, position = "left", font_size = 14) %>%
  column_spec(1, bold = TRUE) %>%   # Make first column (Variable) bold
  column_spec(2, width = "5cm") %>% # Adjust HR column width
  column_spec(3, width = "6cm") %>% # Adjust CI column width
  column_spec(4, width = "4cm") %>% # Adjust P-value column width
  footnote(general = "Source: Computed results based on multivariate Cox PH model.")


```{r,Spatial Analysis}
file.exists("C:/Users/pc/Desktop/Counties Shape File/ke_subcounty.shp")
sub_county <- sub_county <- st_read("C:\\Users\\pc\\Desktop\\ke_subcounty.shp")
lake_victoria <- st_read("C:\\Users\\pc\\Desktop\\Lake_Victoria_Shapefile.shp")
#facilities <- st_read("C:\\Users\\pc\\Desktop\\homabay\\Homabay health Facilities shapefiles.shp")

print(sub_county)
ggplot(data = sub_county) + geom_sf(fill = "lightblue", color = "black", size = 0.3) + labs(title = "Kenya Sub-Counties Map") + theme_minimal()

```

<!-- Plotting Homa Bay-->
```{r,Mapping Homa Bay SU cOUNTIES}

homabay_shp <- sub_county %>% filter(county == "Homa Bay") 

ggplot() + geom_sf(data = homabay_shp, fill = "lightblue", color = "black") + ggtitle("Map of Homa Bay Sub-Counties") + theme_minimal()
```



```{r, Hiv Cases Per Sub County}
status_counts <- df %>%
  group_by(subcounty) %>%
  summarize(monthsantibodytest = n())



status_counts1 <- status_counts %>%
  mutate(subcounty = case_when(
    subcounty == "Homa Bay" ~ "Homa Bay Town Sub County",
    subcounty == "Karachauony" ~ "Karachuonyo Sub County",
    subcounty == "Ragwe" ~ "Rangwe Sub County",
    subcounty == "Suba" ~ "Suba Sub County",
    subcounty == "Mbita" ~ "Mbita Sub County",
    subcounty == "Kasipul" ~ "Kasipul Sub County",
    subcounty == "Kabondo Kasipul" ~ "Kabondo Kasipul Sub County",
    subcounty == "Ndhiwa" ~ "Ndhiwa Sub County",
    TRUE ~ subcounty # Keep other names unchanged
  ))

homabay_shp <- homabay_shp %>%
  left_join(status_counts1, by = "subcounty")
  
homabay_shp %>%
  select(subcounty, monthsantibodytest) %>%
  st_drop_geometry() %>%
  head(10) %>%
  kable(caption = "Total HIV Cases per Subcounty in Homa Bay") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed", "responsive"))


ggplot(data = homabay_shp) +
  geom_sf(aes(fill = monthsantibodytest), color = "white", size = 0.3) +  # Map outline
  scale_fill_viridis_c(option = "magma", direction = -1, na.value = "grey80", name = "Total Cases") +  
  geom_sf_text(aes(label = subcounty), size = 3, color = "white") +  # Subcounty labels
  labs(
    title = "Geographical Distribution of HIV Cases in Homa Bay",
    subtitle = "Sub-county level analysis",
    x = "",
    y = "",
    caption = "Data Source: [Your Source Here]"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    panel.background = element_rect(fill = "lightgrey"),  # Light grey background for contrast
    panel.grid = element_blank()  # Remove grid for clarity
  )
```

#patient

patients_count <- df1 %>%
  group_by(subcounty) %>%
  summarize(count = n()) %>% 
  rename(`Number of Patients` = count) 
 

patient_count1 <- patients_count %>%
  mutate(subcounty = case_when(
    subcounty == "Homa Bay" ~ "Homa Bay Town Sub County",
    subcounty == "Karachauony" ~ "Karachuonyo Sub County",
    subcounty == "Ragwe" ~ "Rangwe Sub County",
    subcounty == "Suba" ~ "Suba Sub County",
    subcounty == "Mbita" ~ "Mbita Sub County",
    subcounty == "Kasipul" ~ "Kasipul Sub County",
    subcounty == "Kabondo Kasipul" ~ "Kabondo Kasipul Sub County",
    subcounty == "Ndhiwa" ~ "Ndhiwa Sub County",
    TRUE ~ subcounty # Keep other names unchanged
  ))


homabay_shp3 <- homabay_shp %>%
  left_join(patient_count1, by = "subcounty")
  
 
homabay_shp3 %>%
  select(subcounty, `Number of Patients`) %>%
  st_drop_geometry() %>%
  head(10) %>%
  kable(caption = "Total Patients per Subcounty in Homa Bay") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed", "responsive"))


ggplot(data = homabay_shp3) +
  geom_sf(aes(fill = `Number of Patients`), color = "white", size = 0.3) +  # Map outline
  scale_fill_gradient(low = "lightblue", high = "darkblue", na.value = "grey80", name = "Total Patients") +  
  geom_sf_text(aes(label = subcounty), size = 3, color = "white") +  # Subcounty labels
  labs(
    title = "Geographical Distribution of HIV Cases in Homa Bay",
    subtitle = "Sub-county level analysis",
    x = "",
    y = "",
    caption = "Data Source: [Your Source Here]"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    panel.background = element_rect(fill = "lightgrey"),  # Light grey background for contrast
    panel.grid = element_blank()  # Remove grid for clarity
  )

  
  
  

subcounty_codes <- as.numeric(as.factor(df1$subcounty))
df1$subcounty_code <- subcounty_codes

str(coords)
dim(coords)
head(coords)

coords <- as.matrix(df1$subcounty_code)
nb <- knn2nb(knearneigh(coords, k = 3))


#reading health facilities
data_facilities <- read_csv("D:/Downloads/Homabay health Facilities.csv")
data_facilities1 <- data.frame(data_facilities)
colnames(data_facilities1)


df1$subcounty <- tolower(df1$subcounty)
data_facilities1$Admin1 <- tolower(data_facilities1$Admin)

merged_df <- df1 %>% left_join(data_facilities, by = c("subcounty"= "Admin1"))


head(merged_df)


colnames(merged_df)
merged_df <- merged_df %>% filter(!is.na(Lat) & !is.na(Long) & !is.na(monthsantibodytest))

merged_df$Lat <- as.numeric(merged_df$Lat)
merged_df$Long <- as.numeric(merged_df$Long)

merged_df$hiv_antigen_positive <- ifelse(merged_df$monthsantibodytest == "positive",1,0)

y = as.numeric(merged_df$hiv_antigen_positive)


coords <- as.matrix(merged_df[,c("Long","Lat")])
nb <- nb <- knn2nb(knearneigh(coords, k = 3))



```{r Create spatial weights}

lw <- nb2listw(nb, style = "W")

```

```{r Compute Moran's I}

moran_test <- moran.test(y, lw)
```
```{r}
print(moran.test)
```

```{r Convert data to spatial format for mapping}

merged_sf <- st_as_sf(merged_df, coords = c("Long", "Lat"), crs = 4326)

```

```{r  Plot density map of HIV antigen positivity}

tmap_mode("view")
tm_shape(merged_sf) + 
  tm_dots(col = "hiv_antigen_positivity", palette = "Reds", size = 0.1, title = "HIV Antigen Positivity") +
  tm_layout(title = "Density Map of Spatial Autocorrelation - Homa Bay")
```





```{r, machine learning}
```

# Convert target variable to factor
df1$monthsantibodytest <- as.factor(df1$monthsantibodytest)

# Define independent variables
independent_vars <- c("age", "weightkg", "bwtkgs", "babywt6wks", "maritalstatus",
                      "residence", "employmentstatus", "ancattendance", "facilitylevelanc",
                      "educationlevel", "adherence")

# Convert categorical variables to factors (excluding numeric ones)
categorical_vars <- c("maritalstatus", "residence", "employmentstatus", 
                      "ancattendance", "facilitylevelanc", "educationlevel", "adherence")

df1[categorical_vars] <- lapply(df1[categorical_vars], as.factor)

# Ensure numeric variables are correctly typed
numeric_vars <- c("age", "weightkg", "bwtkgs", "babywt6wks")
df1[numeric_vars] <- lapply(df1[numeric_vars], as.numeric)

# Split data into training and testing sets
set.seed(123)
trainIndex <- createDataPartition(df1$monthsantibodytest, p = 0.8, list = FALSE)
trainData <- df1[trainIndex, ]
testData <- df1[-trainIndex, ]  # Fixed: Corrected from 'data' to 'df1'

# Ensure categorical variables in testData match levels in trainData
for (var in categorical_vars) {
  testData[[var]] <- factor(testData[[var]], levels = levels(trainData[[var]]))
}

# Ensure numeric variables remain numeric
testData[numeric_vars] <- lapply(testData[numeric_vars], as.numeric)

# Train logistic regression model
model <- glm(monthsantibodytest ~ ., 
             data = trainData[, c(independent_vars, "monthsantibodytest")], 
             family = binomial)

# Summary of the model
summary(model)

# Predict on the test set
predictions <- predict(model, newdata = testData[, independent_vars], type = "response")

# Print first few predictions
head(predictions)

class_predictions <- ifelse(predictions >= 0.5, 1, 0)
head(class_predictions)

# Convert actual test labels to numeric
actual_values <- as.numeric(as.character(testData$monthsantibodytest))

# Create a confusion matrix
conf_matrix <- confusionMatrix(as.factor(class_predictions), as.factor(actual_values))

# Print accuracy
accuracy <- conf_matrix$overall["Accuracy"]
print(paste("Model Accuracy:", round(accuracy, 4)))


# Create a data frame with actual vs predicted values
plot_data <- data.frame(
  Actual = actual_values,
  Predicted_Prob = predictions
)

# Plot actual vs predicted values
ggplot(plot_data, aes(x = Actual, y = Predicted_Prob, color = Predicted_Prob > 0.8)) +
  geom_jitter(width = 0.2, height = 0.02, alpha = 0.7, size = 2) +
  scale_color_manual(values = c("FALSE" = "blue", "TRUE" = "red")) +
  labs(title = "Test vs Predicted Probabilities", x = "Actual Values", y = "Predicted Probability") +
  theme_minimal() +
  theme(legend.position = "none")




