---
title: "HIV/AIDS Modelling"
author: "Bonsoul"
date: "2025-02-26"
output: html_document
---

install.packages("maditr")
install.packages("expss")

```{r setup, include=TRUE , warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)  

library(rnaturalearth)
library(rnaturalearthdata)
library(CARBayes)
library(gridExtra)             
library(grid) 
library(survival)
library(tidyverse) 
library(caret) 
library(randomForest) 
library(ggplot2) 
library(dplyr) 
library(corrplot) 
library(readxl) 
library(tidyr) 
library(survminer) 
library(sf) 
library(tmap)
library(RColorBrewer)
library(knitr)
library(mgcv)
library(knitr)
library(kableExtra)
library(broom)
library(spdep)
library(brglm2)

```

```{r, Readind the data , warning=FALSE}

df <- read_excel("D:/Downloads/HIV - VARIABLES.xlsx")

str(df)
```

<!--Rename coordinate columns for clarity-->

df <- df %>% rename(latitude = var52, longitude = var51)

<!--Summary of the dataset to verify changes-->

summary(df)

<!--Data Exploration-->

duplicates <- df[duplicated(df),]

print(duplicates)

<!--No duplicates-->

missing_per_column <- colSums(is.na(df)) 
total_missing <- sum(missing_per_column) 
print(total_missing) 
print(missing_per_column)

columns_with_missing <- names(missing_per_column[missing_per_column > 0]) 
print(columns_with_missing)

<!--Compute summary statistics for numeric columns-->

summary_stats <- df %>% 
  select(where(is.numeric)) %>% 
  summarise(across(
    everything(), 
    list(
      min = ~min(., na.rm = TRUE),
      Q1 = ~quantile(., 0.25, na.rm = TRUE),
      median = ~median(., na.rm = TRUE),
      mean = ~mean(., na.rm = TRUE),
      Q3 = ~quantile(., 0.75, na.rm = TRUE),
      max = ~max(., na.rm = TRUE),
      sd = ~sd(., na.rm = TRUE)
    )
  ))


summary(summary_stats)

<!--Visualize distributions with histograms-->

numeric_cols <- df %>% select(where(is.numeric)) %>% colnames() 
numeric_cols <- setdiff(numeric_cols, c("longitude", "latitude"))

for (col in numeric_cols) { 
  p <- ggplot(df, aes_string(x = col)) + 
    geom_histogram(binwidth = 30, fill = "blue", alpha = 0.7) + 
    ggtitle(paste("Histogram of", col)) + 
    theme_minimal() 
  print(p)}


png('histogram.png') dev.off()

<!--Boxplot-->

for (col in numeric_cols) { 
p <- ggplot(df, aes(y = .data[[col]])) +  
    geom_boxplot(fill = "lightblue", na.rm = TRUE) + 
    ggtitle(paste("Boxplot of", col)) + 
    theme_minimal() 
    print(p)  
}


#Mutating +ve and -ne to be numerical

df1 <- df %>% mutate(monthsantibodytest = ifelse(tolower(monthsantibodytest) == "positive", 1, 0))

table(df1$monthsantibodytest)

colnames(df1)


<!--Create survival object-->

surv_obj <- Surv(time = df1$time, event = df1$monthsantibodytest)

<!--Fit Kaplan-Meier survival model-->

km_fit <- survfit(surv_obj ~ 1, data = df1)

<!--Plot Kaplan-Meier curve-->

ggsurvplot(km_fit, data = df1, conf.int = TRUE, ggtheme = theme_minimal(), title = "Kaplan-Meier Survival Curve", xlab = "Time (Months)", ylab = "HIV-Free Survival Probability")

```{=html}
<!--Chi-square test used for comparisons
-->
```

colnames(df1)

dependent_var <- "monthsantibodytest"

categorical_vars <- c(
  "subcounty", "maritalstatus", "historyofstiduringpregnancy", "hypertension", 
  "partnershivstatus", "whohivdiseasestage", "t_membrane_r", "sexofthebaby", 
  "residence", "syphillis", "cd4cellcountcellsmm3", "employmentstatus", 
  "ancattendance", "facilitylevelanc", "hivstatusbeforepregnancy", "syphilis", 
  "uti", "miscarriage", "abortion", "stillbirth", "premature", 
  "haartduringpregnancy", "cd4cellcount", "historysti_during_pregnancy", "distancetofacility",
  "genitalwart", "vaginaldischargesyndrome", "genitalherpes", 
  "treatedduringpregnancy", "malaria", "anaemia", "hypertension", 
  "partnerhivstatus", "whohivdiseasestage", "t_membrane_r", "sexbaby", 
  "educationlevel", "adherence","tmembraner", "hypertention","patnershivstatus"
)



all_vars <- categorical_vars[categorical_vars %in% names(df1)]


chi_results <- lapply(all_vars, function(var) {
  table <- table(df[[var]], df[[dependent_var]])
  test <- chisq.test(table)
  data.frame(variable = var, chi2_value = test$statistic, p_value = test$p.value)
})

# Combine the results into a single data frame
chi_result_df <- do.call(rbind, chi_results)

kable(chi_result_df, format = "html", digits = 4, caption = "Table 1: Chi-square Test Results for HIV Antigen Positivity") %>%
  kable_styling(full_width = FALSE, position = "left", font_size = 14) %>%
  column_spec(1, bold = TRUE) %>%   
  column_spec(2, width = "5cm") %>% 
  column_spec(3, width = "5cm") %>% 
  footnote(general = "Source: Computed results based on chi-square test.")





<!--  Factors Associated with HIV Antigen Positivity by Bivariate Logistic Regression-->


dependent_var <- "monthsantibodytest"  

# List of independent variables for bivariate logistic regression
independent_vars <- c(
  "ancattendance","haartduringpregnancy", "whohivdiseasestage","numberofancvisitsmade","lduration","durationofbfmonths","educationlevel","adherence","hivstatusbeforepregnancy","premature") 

# Filter variables that exist in df
available_vars <- independent_vars[independent_vars %in% names(df1)]



# Function to perform bivariate logistic regression
logit_results <- lapply(available_vars, function(var) {
  formula <- as.formula(paste(dependent_var, "~", var))
  model <- glm(formula, data = df1, family = binomial)
  tidy_model <- tidy(model, conf.int = TRUE)  # Get summary with confidence intervals
  
  # Format output
  data.frame(
    Variable = var,
    Odds_Ratio = round(exp(tidy_model$estimate[2]), 4),  # Exponentiate coefficient
    CI = paste0("[", round(exp(tidy_model$conf.low[2]), 3), "-", round(exp(tidy_model$conf.high[2]), 3), "]"),
    P_Value = round(tidy_model$p.value[2], 4)
  )
})

bivar_logit_results_df <- do.call(rbind, logit_results)


kable(bivar_logit_results_df, format = "html", digits = 4, caption = "Table 4.2: Factors Associated with HIV Antigen Positivity by Bivariate Logistic Regression") %>%
  kable_styling(full_width = FALSE, position = "left", font_size = 14) %>%
  column_spec(1, bold = TRUE) %>%   
  column_spec(2, width = "5cm") %>% 
  column_spec(3, width = "5cm") %>% 
  column_spec(4, width = "5cm") %>% 
  footnote(general = "Source: Computed results based on bivariate logistic regression.")



<!--Factors Associated With HIV Antigen Positivity by Multivariate Logistic Regression-->

# Define and run the multivariate logistic regression model
multivar_model <- glm(
  monthsantibodytest ~ lduration + durationofbfmonths + 
    ancattendance + haartduringpregnancy + hivstatusbeforepregnancy + educationlevel + 
    adherence + numberofancvisitsmade + premature,
  data = df1, 
  family = binomial
)

# Extract results
multivar_results <- tidy(multivar_model, conf.int = TRUE) %>%
  filter(term != "(Intercept)") %>%
  mutate(
    Variable = recode(term, 
                      "duration_labor" = "Duration of labor",
                      "duration_breastfeeding" = "Duration of breast feeding in months",
                      "antenatal_clinic_attendance" = "Antenatal clinic attendance",
                      "haart_during_pregnancy" = "Initiation of HAART during pregnancy",
                      "never_attended_school" = "Never attended school",
                      "attended_primary" = "Attended up to Primary school",
                      "attended_secondary" = "Attended up to Secondary school",
                      "adherencepoor" = "Adherence to HAART during ANC (Poor)",
                      "adherencegood" = "Adherence to HAART during ANC (Good)",
                      "no_anc_visits" = "No of ANC visits made during pregnancy",
                      "premature" = "Premature"),
    Odds_Ratio = round(exp(estimate), 4),
    CI = paste0("[", round(exp(conf.low), 3), "-", round(exp(conf.high), 3), "]"),
    P_Value = round(p.value, 4)
  ) %>%
  select(Variable, Odds_Ratio, CI, P_Value)

# Display the results in a table

kable(multivar_results, format = "html", digits = 4, 
      caption = "Table 4.3: Factors Associated With HIV Antigen Positivity by Multivariate Logistic Regression") %>%
  kable_styling(full_width = FALSE, position = "left", font_size = 14) %>%
  column_spec(1, bold = TRUE) %>%  
  column_spec(2:4, width = "5cm") %>%
  footnote(general = "Source: Computed results based on multivariate logistic regression analysis.")



<!--Bivariate Cox PH Model-->

# List of independent variables
independent_vars <- c("employmentstatus", "ancattendance", "facilitylevelanc",
                      "hivstatusbeforepregnancy", "premature","haartduringpregnancy",
                      "historyofstiduringpregnancy", "cd4cellcountcellsmm3", "whohivdiseasestage",
                      "numberofancvisitsmade", "durationofbfmonths")

# Apply Bivariate Cox PH Models for each variable

cox_results <- lapply(independent_vars, function(var) {
  formula <- as.formula(paste("Surv(time, monthsantibodytest) ~", var))
  model <- coxph(formula, data = df1)
  tidy_model <- tidy(model, conf.int = TRUE)  # Extract HR, CI, p-value
  
  # Format the output
  data.frame(
    Variable = var,
    HR = round(exp(tidy_model$estimate), 3),  # Exponentiate to get Hazard Ratio
    CI = paste0("[", round(exp(tidy_model$conf.low), 3), "-", round(exp(tidy_model$conf.high), 3), "]"),
    P_Value = round(tidy_model$p.value, 3)
  )
})

# Combine results into a single table
cox_results_df <- bind_rows(cox_results)


# Generate Harvard-style table with kableExtra
kable(cox_results_df, format = "html", digits = 4, caption = "Table 4.4: Factors Associated With HIV Antigen Positivity by Bivariate Cox PH Model") %>%
  kable_styling(full_width = FALSE, position = "left", font_size = 14) %>%
  column_spec(1, bold = TRUE) %>%   # Make first column (Variable) bold
  column_spec(2, width = "5cm") %>% # Adjust HR column width
  column_spec(3, width = "6cm") %>% # Adjust CI column width
  column_spec(4, width = "4cm") %>% # Adjust P-value column width
  footnote(general = "Source: Computed results based on statistical analysis.")

<!--Multivariate-->

colnames(df1)

# List of independent variables
independent_vars <- c("hivstatusbeforepregnancy","haartduringpregnancy","cd4cellcountcellsmm3", "whohivdiseasestage","adherence")

# Create a formula with all independent variables for the multivariate model
multivariate_formula <- as.formula(paste("Surv(time, monthsantibodytest) ~", paste(independent_vars, collapse = " + ")))

# Apply Multivariate Cox PH Model
multivariate_model <- coxph(multivariate_formula, data = df1)

# Tidy the model to extract HR, CI, and p-values
tidy_multivariate_model <- tidy(multivariate_model, conf.int = TRUE)

# Format the output
multivariate_results_df <- data.frame(
  Variable = tidy_multivariate_model$term,
  HR = round(exp(tidy_multivariate_model$estimate), 3),  # Exponentiate to get Hazard Ratio
  CI = paste0("[", round(exp(tidy_multivariate_model$conf.low), 3), "-", round(exp(tidy_multivariate_model$conf.high), 3), "]"),
  P_Value = round(tidy_multivariate_model$p.value, 3)
)

# Generate Harvard-style table with kableExtra
kable(multivariate_results_df, format = "html", digits = 4, caption = "Table 4.X: Factors Associated With HIV Antigen Positivity by Multivariate Cox PH Model") %>%
  kable_styling(full_width = FALSE, position = "left", font_size = 14) %>%
  column_spec(1, bold = TRUE) %>%   # Make first column (Variable) bold
  column_spec(2, width = "5cm") %>% # Adjust HR column width
  column_spec(3, width = "6cm") %>% # Adjust CI column width
  column_spec(4, width = "4cm") %>% # Adjust P-value column width
  footnote(general = "Source: Computed results based on multivariate Cox PH model.")


<!--Spatial Analysis-->

```{r Spatial analysis}
file.exists("C:/Users/pc/Desktop/Counties Shape File/ke_subcounty.shp")
sub_county <- sub_county <- st_read("C:\\Users\\pc\\Desktop\\ke_subcounty.shp")
lake_victoria <- st_read("C:\\Users\\pc\\Desktop\\Lake_Victoria_Shapefile.shp")
#facilities <- st_read("C:\\Users\\pc\\Desktop\\homabay\\Homabay health Facilities shapefiles.shp")

print(sub_county)
ggplot(data = sub_county) + geom_sf(fill = "lightblue", color = "black", size = 0.3) + labs(title = "Kenya Sub-Counties Map") + theme_minimal()

```

<!-- Plotting Homa Bay-->
```{r,Mapping Homa Bay SU cOUNTIES}

homabay_shp <- sub_county %>% filter(county == "Homa Bay") 

ggplot() + geom_sf(data = homabay_shp, fill = "lightblue", color = "black") + ggtitle("Map of Homa Bay Sub-Counties") + theme_minimal()
```



<!-- Hiv Cases Per Sub County -->

status_counts <- df %>%
  group_by(subcounty) %>%
  summarize(monthsantibodytest = n())
view(status_counts)


status_counts1 <- status_counts %>%
  mutate(subcounty = case_when(
    subcounty == "Homa Bay" ~ "Homa Bay Town Sub County",
    subcounty == "Karachauony" ~ "Karachuonyo Sub County",
    subcounty == "Ragwe" ~ "Rangwe Sub County",
    subcounty == "Suba" ~ "Suba Sub County",
    subcounty == "Mbita" ~ "Mbita Sub County",
    subcounty == "Kasipul" ~ "Kasipul Sub County",
    subcounty == "Kabondo Kasipul" ~ "Kabondo Kasipul Sub County",
    subcounty == "Ndhiwa" ~ "Ndhiwa Sub County",
    TRUE ~ subcounty # Keep other names unchanged
  ))




homabay_shp <- homabay_shp %>%
  left_join(status_counts1, by = "subcounty")
  
  
homabay_shp %>%
  select(subcounty, monthsantibodytest.y.y) %>%
  st_drop_geometry() %>%
  head(10) %>%
  kable(caption = "Total HIV Cases per Subcounty in Homa Bay") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed", "responsive"))


ggplot(data = homabay_shp) +
  geom_sf(aes(fill = monthsantibodytest.y.y), color = "white", size = 0.3) +  # Map outline
  scale_fill_viridis_c(option = "magma", direction = -1, na.value = "grey80", name = "Total Cases") +  
  geom_sf_text(aes(label = subcounty), size = 3, color = "white") +  # Subcounty labels
  labs(
    title = "Geographical Distribution of HIV Cases in Homa Bay",
    subtitle = "Sub-county level analysis",
    x = "",
    y = "",
    caption = "Data Source: [Your Source Here]"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    panel.background = element_rect(fill = "lightgrey"),  # Light grey background for contrast
    panel.grid = element_blank()  # Remove grid for clarity
  )










<!--patient-->

patients_count <- df1 %>%
  group_by(subcounty) %>%
  summarize(count = n()) %>% 
  rename(`Number of Patients` = count) 
 

view(patients_count)

colnames(patients_count)

patient_count1 <- patients_count %>%
  mutate(subcounty = case_when(
    subcounty == "Homa Bay" ~ "Homa Bay Town Sub County",
    subcounty == "Karachauony" ~ "Karachuonyo Sub County",
    subcounty == "Ragwe" ~ "Rangwe Sub County",
    subcounty == "Suba" ~ "Suba Sub County",
    subcounty == "Mbita" ~ "Mbita Sub County",
    subcounty == "Kasipul" ~ "Kasipul Sub County",
    subcounty == "Kabondo Kasipul" ~ "Kabondo Kasipul Sub County",
    subcounty == "Ndhiwa" ~ "Ndhiwa Sub County",
    TRUE ~ subcounty # Keep other names unchanged
  ))


homabay_shp3 <- homabay_shp %>%
  left_join(patient_count1, by = "subcounty")
  
 
homabay_shp3 %>%
  select(subcounty, `Number of Patients`) %>%
  st_drop_geometry() %>%
  head(10) %>%
  kable(caption = "Total Patients per Subcounty in Homa Bay") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed", "responsive"))


ggplot(data = homabay_shp3) +
  geom_sf(aes(fill = `Number of Patients`), color = "white", size = 0.3) +  # Map outline
  scale_fill_gradient(low = "lightblue", high = "darkblue", na.value = "grey80", name = "Total Patients") +  
  geom_sf_text(aes(label = subcounty), size = 3, color = "white") +  # Subcounty labels
  labs(
    title = "Geographical Distribution of HIV Cases in Homa Bay",
    subtitle = "Sub-county level analysis",
    x = "",
    y = "",
    caption = "Data Source: [Your Source Here]"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    panel.background = element_rect(fill = "lightgrey"),  # Light grey background for contrast
    panel.grid = element_blank()  # Remove grid for clarity
  )

  

  
  
  
